---
title: "Geom Bar in R"
author: "Nicole Fimmen"
date: '2018-07-26'
slug: geom-bar-in-r
tags: []
categories: R
---


https://www.sharpsightlabs.com/blog/ultimate-guide-bar-chart-geom_bar/


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# Geom Bar with ggplot2

We're going to practice the basic bar chart with ggplot2. We will be using the data set Arthritis from the vcd package. 

Full disclosure, I'm doing the exact excercise laid out on Sharp Sight Data <https://www.sharpsightlabs.com/blog/ultimate-guide-bar-chart-geom_bar/>.

Make sure you have the vcd library downloaded. Here are the libraries we will be using.  

```{r set up}

library(tidyverse)
library(forcats)
library(vcd)

ggplot(data = Arthritis, aes(x = Sex)) +
  geom_bar()

```

Let's checkout what all is in the Athritis dataframe. It is very important to inspect our data to understand what we are working with. Using the head function will do what it sounds like and gives you the first few rows of the data. 

Glimpse goes into how many variables and observations there are. 

``` {r Arthritis dataframe}
head(Arthritis)

glimpse(Arthritis)
```

# ggplot Building Blocks

A good way to think of ggplot is like building blocks. It follows a systemtic flow each time you create a visualization. 

1. function - ggplot()  
2. data parameter and this time it was "Arthritis" dataframe. 
3. function - aes() this maps the variable to the x axis.
4. geom - this is the object that we'r drawing, whether a bar graph, scatterplot or whatever. 

``` {r geom bar}
ggplot(data = Arthritis, aes(x = Sex)) +
  geom_bar()

```

# Modify
Now we will do two simply changes.
1. change interior color of the bars
2. change the line color of the bars

In order to change the color, we will use the fill aesthetic, not the color one. 

```{r}
ggplot(data = Arthritis, aes(x = Sex)) +
  geom_bar(fill = 'cyan')
```

Now lets try changing the border.

```{r}
ggplot(data = Arthritis, aes(x = Sex)) +
  geom_bar(color = 'red')

```


Not the most intuitive but remember fill is for filling in the bar and color is for the outline color. 

Time to modify the bar chart even more.

1. Horizontal Bar Chart
This is a simple task. Simply add coord_flip() to the end of the code. 

```{r}
ggplot(data = Arthritis, aes(x = Sex)) + geom_bar() + coord_flip()

```


2. Stacked Bar Chart
This is simple also. Map a categorical variable to the fill asethetic. 

```{r}
ggplot(data = Arthritis, aes(x = Treatment)) +
  geom_bar(aes(fill = Improved))

```
3. 100% Stacked Bar Chart
Here instead of the bar length mapping to the number of records, it will fill in the whole axis. 


```{r}

ggplot(data = Arthritis, aes(x = Treatment)) +
  geom_bar(aes(fill = Improved), position = "fill")


```

# Including Plots

You can also embed plots. See Figure \@ref(fig:pie) for example:

```{r pie, fig.cap='A fancy pie chart.', tidy=FALSE}
par(mar = c(0, 1, 0, 1))
pie(
  c(280, 60, 20),
  c('Sky', 'Sunny side of pyramid', 'Shady side of pyramid'),
  col = c('#0292D8', '#F7EA39', '#C4B632'),
  init.angle = -50, border = NA
)
```