---
title: Geom Bar in R
author: Nicole Fimmen
date: '2018-07-26'
slug: geom-bar-in-r
categories:
  - R
tags: []
---


https://www.sharpsightlabs.com/blog/ultimate-guide-bar-chart-geom_bar/


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# Geom Bar with ggplot2

We're going to practice the basic bar chart with ggplot2. We will be using the data set Arthritis from the vcd package. 

Full disclosure, I'm doing the exact excercise laid out on Sharp Sight Data <https://www.sharpsightlabs.com/blog/ultimate-guide-bar-chart-geom_bar/>.

Make sure you have the vcd library downloaded. Here are the libraries we will be using.  

```{r set up}

library(tidyverse)
library(forcats)
library(vcd)

ggplot(data = Arthritis, aes(x = Sex)) +
  geom_bar()

```

# Including Plots

You can also embed plots. See Figure \@ref(fig:pie) for example:

```{r pie, fig.cap='A fancy pie chart.', tidy=FALSE}
par(mar = c(0, 1, 0, 1))
pie(
  c(280, 60, 20),
  c('Sky', 'Sunny side of pyramid', 'Shady side of pyramid'),
  col = c('#0292D8', '#F7EA39', '#C4B632'),
  init.angle = -50, border = NA
)
```